---
title: "Introduction to ggvis"
author: "Sanatan Das"
date: "April 25, 2018"
output:
  ioslides_presentation: default
  beamer_presentation: default
  slidy_presentation: default
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(ggvis)
library(shiny)
library(knitr)

```

## Before We Start on ggvis

Let's talk about ggplot2

- Its the Grammar of Graphics
- Creates in 2005 by Hadley Wickham
- Well Documented
- In 2014, shifts to maintenance mode (no more development)


## What is ggvis?

It is a synthesis of ideas:

- Created in 2014 by Winston Chang
- Grammar of Graphics (ggplot2)
- Data Pipeline and Data Transformation (dplyr)
- Reactivity and Interactivity (shiny)
- Web (vega.js)


## Grammar of Graphics

```{r var1, echo=TRUE, eval=TRUE}
mtcars %>% # data
  ggvis(x = ~wt, y = ~mpg) %>% # axis
  layer_points() %>% # layer1
  layer_smooths() # layer2

```

## Grammar of Graphics

```{r var2, echo=TRUE, eval=TRUE}
mtcars %>% # data
  ggvis(x = ~wt, y = ~mpg, fill = ~factor(cyl)) %>% 
  # axis + property mapping
  layer_points()# layer1

```

## Grammar of Graphics

```{r var3, echo=TRUE, eval=TRUE}
mtcars %>% 
  ggvis(~mpg, ~disp, shape = ~factor(cyl)) %>% 
  layer_points()

```

## Grammar of Graphics

```{r var4, echo=TRUE, eval=TRUE}
n <- 10000
gamma_samp <- tibble(x = rgamma(n, shape = 1, scale = 2))
gamma_samp %>% # data
  # axis + property mapping
  ggvis(x = ~x) %>% 
  layer_histograms()# layer1

```

## Grammar of Graphics

**Layers**

- Simple layers : *layer_points(), layer_paths(), layer_ribbons(), layer_rects(), layer_text()*

- Compound layers : *layer_lines(), layer_histograms(), layer_freqpolys(), layer_smooths()* 

## Examples

```{r varN1, echo=TRUE, eval=TRUE}

df <- data.frame(x = 1:10, y = runif(10))
df %>% ggvis(~x, ~y) %>% layer_paths()

```

## Examples

```{r varN2, echo=TRUE, eval=TRUE}

t <- seq(0, 2 * pi, length = 100)
df <- data.frame(x = sin(t), y = cos(t))
df %>% ggvis(~x, ~y) %>% layer_paths(fill := "red")

```

## Examples

```{r varN3, echo=TRUE, eval=TRUE}

df <- data.frame(x = 1:10, y = runif(10))
df %>% ggvis(~x, ~y) %>% layer_ribbons()

```

## Examples

```{r varN4, echo=TRUE, eval=TRUE}

df %>% ggvis(~x, ~y + 0.1, y2 = ~y - 0.1) %>% layer_ribbons()

```

## Examples

```{r varN5, echo=TRUE, eval=TRUE}

set.seed(1014)
df <- data.frame(x1 = runif(5), x2 = runif(5), y1 = runif(5), y2 = runif(5))
df %>% ggvis(~x1, ~y1, x2 = ~x2, y2 = ~y2, fillOpacity := 0.1) %>% layer_rects()

```

## Examples

```{r varN6, echo=TRUE, eval=TRUE}

df <- data.frame(x = 3:1, y = c(1, 3, 2), label = c("a", "b", "c"))
df %>% ggvis(~x, ~y, text := ~label) %>% layer_text()

```

## Examples

```{r varN7, echo=TRUE, eval=TRUE}

df %>% ggvis(~x, ~y, text := ~label) %>% layer_text(fontSize := 50)

```

## Data Pipeline

Each ggvis function takes a visualization object as an input and returns a modified visualization object as output.

```{r var5, echo=TRUE, eval=TRUE}
mtcars %>% # data
  ggvis(x = ~wt, y = ~mpg) %>% # create ggvis object
  layer_points() %>% # passit to next layer
  layer_smooths()

```

## Data Transformation

```{r var6, echo=TRUE, eval=TRUE}
mtcars %>%
  ggvis(x = ~mpg, y = ~disp) %>%
  # convert engine displacment to litres
  mutate(disp = disp / 61.0237) %>% 
  layer_points()

```

## Reactivity and Interactivity

**How it works?**

- In "regular" programming, function call happens once. The function takes an input and returns an output.

- In functional reactive programming, a reactive can use a value from another reactive; this creates a dependency graph of reactives. The reactives persist.

- When the value of an ancestor node changes, it triggers a recomputation of all its descendants.


## Reactivity and Interactivity

```{r var7, echo=TRUE, eval=TRUE}
n <- 10000
gamma_samp <- tibble(x = rgamma(n, shape = 1, scale = 2))
gamma_samp %>% # data
  ggvis(x = ~x) %>% # axis + property mapping
  layer_histograms(width = 
                     input_slider(min = 1, max = 20, value = 11))# layer1

```

## Reactivity and Interactivity

```{r var8, echo=TRUE, eval=TRUE}

mtcars %>% 
  ggvis(~wt, ~mpg, 
    size := input_slider(10, 100),
    opacity := input_slider(0, 1)
  ) %>% 
  layer_points()

```

## Web

```{r var9, echo=TRUE, eval=TRUE}

## Run these examples only in interactive R sessions
if (interactive()) {

all_values <- function(x) {
  if(is.null(x)) return(NULL)
  paste0(names(x), ": ", format(x), collapse = "<br />")
}

base <- mtcars %>% ggvis(x = ~wt, y = ~mpg) %>%
  layer_points()
base %>% add_tooltip(all_values, "hover")
base %>% add_tooltip(all_values, "click")

# The data sent from client to the server 
# contains only the data columns that
# are used in the plot. If you want to get 
# other columns of data, you should
# to use a key to line up the item 
# from the plot with a row in the data.
mtc <- mtcars
mtc$id <- 1:nrow(mtc)  # Add an id column to use ask the key

all_values <- function(x) {
  if(is.null(x)) return(NULL)
  row <- mtc[mtc$id == x$id, ]
  paste0(names(row), ": ", format(row), collapse = "<br />")
}

mtc %>% ggvis(x = ~wt, y = ~mpg, key := ~id) %>%
  layer_points() %>%
  add_tooltip(all_values, "hover")

}

```

## The Future

The package is still in development state. Think before moving it to production.

Upcoming features :-

- faceting
- zooming and panning
- performance improvement etc...

